---
title: "data_analysis_project"
author: "Jon Donville,Varun Shankar, Mythili Devarakonda"
date: "16/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
library(janitor)
library(tidycensus)
library(stringr)
library(lubridate)

```


## Read in data 

```{r}
marvel_data <- read_csv("data/marvel-wikia-data.csv")
dc_data <- read_csv("data/dc-wikia-data.csv")
```

## Exploratory analysis

1) Marvel Data

The marvel_data dataset consists of 16,376 rows and 13 columns of data on Marvel comic characters. In such a large dataset, only about 90 comic characters are known to be a "gender or sexual minority", according to the data.

``` {r}

glimpse(marvel_data)

marvel_data %>% filter (! is.na (GSM))

```

2) DC Data

The dc_data dataset consists of 6,896 rows and 13 columns of data on DC comic characters. In such a large dataset, only about 64 comic characters are known to be a "gender or sexual minority", according to the data.

``` {r}

glimpse(dc_data)

dc_data %>% filter (! is.na (GSM))


```
```{r}
# Create an if/else statement to create new columns that group characters into LGBT or not, and neutral color hair or not.

dc_data_lgbt <- dc_data %>%  
mutate(
    lgbt_or_not = case_when(
        GSM == 'Bisexual Characters' ~ "LGBT",
        GSM == 'Homosexual Characters' ~ "LGBT",
      )
  ) %>% 
  mutate(lgbt_or_not = replace_na(lgbt_or_not, "Not LGBT"))
dc_data_lgbt_hair <- dc_data_lgbt %>% 
  mutate(
    neutral_hair_or_not = case_when(
        HAIR == 'Black Hair' ~ "NEUTRAL",
        HAIR == 'Blond Hair' ~ "NEUTRAL",
        HAIR == 'Brown Hair' ~ "NEUTRAL",
        HAIR == 'Auburn Hair' ~ "NEUTRAL", 
        HAIR == 'Bald' ~ "NEUTRAL",
        HAIR == 'Green Hair' ~ "Not",
        HAIR == 'Blue Hair' ~ "Not",
        HAIR == 'Gold Hair' ~ "Not", 
        HAIR == 'Grey Hair' ~ "NEUTRAL",
        HAIR == 'Orange Hair' ~ "Not",
        HAIR == 'Pink Hair' ~ "Not",
        HAIR == 'Platinum Blond Hair' ~ "NEUTRAL", 
        HAIR == 'Purple Hair' ~ "Not",
        HAIR == 'Red Hair' ~ "Not",
        HAIR == 'Reddish Brown Hair' ~ "Neutral",
          HAIR == 'Silver Hair' ~ "Not", 
        HAIR == 'Strawberry Blond Hair' ~ "NEUTRAL",
        HAIR == 'Violet Hair' ~ "Not",
        HAIR == 'White Hair' ~ "Not",
        
      )
  ) 

## Same thing for Marvel Data
marvel_data_lgbt <- marvel_data %>%  
mutate(
    lgbt_or_not = case_when(
        GSM == 'Bisexual Characters' ~ "LGBT",
        GSM == 'Homosexual Characters' ~ "LGBT",
      )
  ) %>% 
  mutate(lgbt_or_not = replace_na(lgbt_or_not, "Not LGBT"))
marvel_data_lgbt_hair <- marvel_data_lgbt %>% 
  mutate(
    neutral_hair_or_not = case_when(
        HAIR == 'Black Hair' ~ "NEUTRAL",
        HAIR == 'Blond Hair' ~ "NEUTRAL",
        HAIR == 'Brown Hair' ~ "NEUTRAL",
        HAIR == 'Auburn Hair' ~ "NEUTRAL", 
        HAIR == 'Bald' ~ "NEUTRAL",
        HAIR == 'Green Hair' ~ "Not",
        HAIR == 'Blue Hair' ~ "Not",
        HAIR == 'Gold Hair' ~ "Not", 
        HAIR == 'Grey Hair' ~ "NEUTRAL",
        HAIR == 'Orange Hair' ~ "Not",
        HAIR == 'Pink Hair' ~ "Not",
        HAIR == 'Platinum Blond Hair' ~ "NEUTRAL", 
        HAIR == 'Purple Hair' ~ "Not",
        HAIR == 'Red Hair' ~ "Not",
        HAIR == 'Reddish Brown Hair' ~ "Neutral",
          HAIR == 'Silver Hair' ~ "Not", 
        HAIR == 'Strawberry Blond Hair' ~ "NEUTRAL",
        HAIR == 'Violet Hair' ~ "Not",
        HAIR == 'White Hair' ~ "Not",
        
      )
  ) 

```