---
title: "lab_02"
author: "derek willis"
date: "8/24/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse and Lubridate. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and lubridate
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with West Virginia loans contained in the PPP loans data.  

**Task** Run the codeblock below to load the data.

```{r}
west_virginia_ppp <- read_rds("data/lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the loan amounts into the following options - under_10k, 10k_50k, 50k_100k, 100k_500k, 500k_1m, 1m_plus - and calculates how many loans are in each group. How would you describe the distribution of loans in a sentence?
Hint: You will need to use `case_when` to make this work.
**A1** [write your answer here]
Most of the loans are under $50,000 and very few are above $500,000.
```{r}
# Write the code to produce your answer here
west_virginia_jobs_categories <- west_virginia_ppp %>%
mutate(
  amount_category = case_when(
    amount < 10000 ~ 'under_10',
    amount >= 10000 & amount < 50000 ~ '10k_50k',
    amount >= 50000 & amount < 100000 ~ '50k_100k',
    amount >= 100000 & amount < 500000 ~ '100k_500k',
    amount >= 500000 & amount < 1000000 ~ '500k_1m',
    amount >= 1000000 ~ '1m_plus'
  )
) %>% group_by(amount_category) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
```

**Q2** Some of the loans appear not to have been paid out - there is a column representing the amount of money that hasn't been disbursed to the applicant. Which servicing lenders have the most loans with undisbursed amounts greater than 0?
**A2** [write your answer here]
Davis Trust company has 12 and Miners & Merchants Bank has 1.
```{r}
# Write the code to produce your answer here
notpaidout <- west_virginia_ppp %>% filter(undisbursed_amount > 0)
notpaidout %>% group_by(lender) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
```

**Q3** What are some potential reasons why a loan might not be paid out? What questions does that circumstance raise for you?
**A3** [write your answer here]
The fact that so many of these undisbursed loans come from one lender makes me think there's an issue with Davis Trust Company in particular and raises questions about whether the issue is on the people applying for the loans through Davis Trust or something about Davis Trust's process that's leading to 12 loans to not be paid off.

**Q4** Which month in 2021 had the most loans approved?
**A4** [write your answer here]
March 2021
```{r}
# Write the code to produce your answer here
west_virginia_ppp %>%
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(Month) %>%
  summarise(total = n()) %>%
  arrange(desc(total)) %>%
  filter(Month > '2021-01-01')

```

**Q5** How many loans went to West Virginia applicants that listed the maximum number of jobs retained (500)? Which one had the earliest approved date?
**A5** [write your answer here]
10, the earliest was approved on April 8, 2020.
```{r}
# Write the code to produce your answer here
maxjobsretained <- west_virginia_ppp %>% filter(jobs_retained == 500)
maxjobsretained %>% group_by(date_approved) %>%
  summarise(total = n()) %>%
  arrange((date_approved))
```
