---
title: "Class Reference"
author: "Student name"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

Jan. 27 Notes
setwd is for setting directories
```{r}
setwd("~/Documents/GitHub/data_journalism_2022_spring")
```

Tidyverse is collection of libraries, opinionated R packages that are designed for dasa science

installing packages 

```{r}
install.packages('tidyverse')
```
```{r}
library(tidyverse)
```
need to do library command to get tidyverse loaded into file

glimpse funciton gives you a view of a lot of the data
head function gives first six data entries
ppp_maryland_loans %>%
  group_by(project_county_name) %>% ##groups by category
  summarise(
    count_loans = n(),
    total_loans_amount = sum(amount) ##gives info about the number of loans by ammount
  ) %>%
  arrange(desc(total_loans_amount))
this is asking the data a question: in this case it's what county has the most loan amounts
  