---
title: "Reverse Engineering Project"
author: "Varun, Mythili, Jonathan"
date: "3/5/2022"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the "CEOs gave heavily during legislative session, exposing loophole in NC's fundraising ban" story, which can be found here https://www.wral.com/in-session-campaign-donations/17716104/


## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(refinr)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


nc_finance_data <- read_csv("data/donors0103-0701.csv")
nc_finance_data <- nc_finance_data %>% mutate(date = mdy(date)) %>% mutate(amount = as.numeric(parse_number(amount)))
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: “Gov. Roy Cooper's main campaign account brought in another $407,000 during session.”
* **Analysis summary**: I tried filtering the data set using his main campaign account "Cooper for North Carolina" but the amount didn't match with the article's result.

```{r}
# Put code to reverse engineer sentence here
# Display results of code below this codeblock
nc_finance_data %>% 
 filter(name == "Cooper For North Carolina")
```  


### Sentence 2

* **Sentence text**: “Berger, R-Rockingham, brought in about $60,000 that week. House Speaker Tim Moore's campaign account collected more than $72,000. Both men declined interview requests on these issues.”

* **Analysis summary**: I was able to recreate both these values by using the right date frame and date format.
```{r}
# Put code to reverse engineer sentence here
berger_donations <- nc_finance_data %>%  
  filter(date >= '2018-05-09' & date <= '2018-05-16' & committee_name == "PHILIP E BERGER COMM") 

berger_donations %>% 
summarize(total_loans_amount = sum(amount))

moore_donations <- nc_finance_data %>%  
  filter(date >= '2018-05-09' & date <= '2018-05-16' & committee_name == "FRIENDS OF TIM MOORE") 

moore_donations %>% 
summarize(total_loans_amount = sum(amount))
```

```{r}
moore_donations <- nc_finance_data %>%  
  filter(date >= "0001-03-18" & date <= "0001-10-18" & committee_name == "FRIENDS OF TIM MOORE") 

moore_donations %>% 
summarize(
      total_loans_amount = sum(amount)
)

```
# Display results of code below this codeblock

### Sentence 3

Sentence text: [R.H. Barringer's board chair, Jasie Barringer, donated another $8,200 in-session, and Teresa Craig kicked in another $5,200 to Berger. The North Carolina Beer & Wine Wholesalers Association PAC gave $18,400 during the week before session and on opening day.]
Analysis summary: [The problem with the data right now is that the amount row is not a number. The datatype will need to be changed before getting any accurate data. I could filter for those names specifically, but it would make it difficult to produce any meaningful obversations other than exactly what I am looking for. ]

```{r}
nc_finance_data %>% 
  group_by("name-clean") %>%
arrange(desc(amount))
```


### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: Senate President Pro Tem Phil Berger's campaign alone raised nearly $222,000 during session. Twenty-three people gave him the maximum allowed: $5,200.”

* **Analysis summary**: By filtering by the date and the committee and then summarizing the total amount of donations, I was able to confirm Berger got just under 222k in donations during the session. 
I then did the same but also added another filter requirement, that the amount donated was the maximum allowed of 5200, and I got 23 results, confirming what the story said.

May 16 to June 29
```{r}
# Put code to reverse engineer sentence here
berger_donations <- nc_finance_data %>%  
  filter(date >= '2018-05-16' & date <= '2018-06-29' & committee_name == "PHILIP E BERGER COMM") 

berger_donations %>% 
summarize(total_loans_amount = sum(amount))

berger_max_donations <- nc_finance_data %>%  
  filter(date >= '2018-05-16' & date <= '2018-06-29' & committee_name == "PHILIP E BERGER COMM" & amount == 5200) 

berger_max_donations
# Display results of code below this codeblock

```

-30-