---
title: "Reverse Engineering Project"
author: "Varun, Mythili, Jonathan"
date: "3/5/2022"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the "CEOs gave heavily during legislative session, exposing loophole in NC's fundraising ban" story, which can be found here https://www.wral.com/in-session-campaign-donations/17716104/


## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(refinr)
library(janitor)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 
nc_finance_data <- read_csv("data/donors0103-0701.csv")
nc_finance_data <- nc_finance_data %>%
  mutate(date = as.Date(date)) %>% 
  mutate(amount = as.numeric(parse_number(amount)))
```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: 

* **Analysis summary**:


### Sentence 2

* **Sentence text**: “Berger, R-Rockingham, brought in about $60,000 that week. House Speaker Tim Moore's campaign account collected more than $72,000. Both men declined interview requests on these issues.”

* **Analysis summary**: I tried to confirm the values found here but ran into a couple of issues. The first was the time-frame, which I think is 1/3 to 1/10 but I'm not sure. The second is knowing where all the donations are going, it's hard to see if I'm catching all of Berger/Moore's donations.
```{r}
# Put code to reverse engineer sentence here
berger_donations <- nc_finance_data %>%  
  filter(date >= "0001-03-18" & date <= "0001-10-18" & committee_name == "PHILIP E BERGER COMM") 

berger_donations %>% 
summarize(
      total_loans_amount = sum(amount)
) 
```

```{r}
moore_donations <- nc_finance_data %>%  
  filter(date >= "0001-03-18" & date <= "0001-10-18" & committee_name == "FRIENDS OF TIM MOORE") 

moore_donations %>% 
summarize(
      total_loans_amount = sum(amount)
)

```
# Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-