---
title: "lab_07"
author: "sean mussenden"
date: "8/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

read html given a url, reads html
html table, given some html, it finds all table objects in it
html element, given some html you supply one or more tags or an xpath expression


To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load rvest, janitor and the tidyverse
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse,rvest,janitor
library(rvest)
library(tidyverse)
library(janitor)
```


Q1. How many individual cases has the U.S. Department of Justice filed against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud?  An example of one case that appears on this page is "U.S. v. Richard Ayvazyan et al". To answer this question, you will need to use rvest to scrape the data on this single webpage into a dataframe that lists only case names. Hint: you will need to make use of html_elements() and html_text() -- a function that extracts text inside of an html tag -- for this.
A1.
There are 101  cases that the DOJ filed against people accused of fraud regarding the PPP/EIDL loan programs.
```{r}
doj_pppfraud_url <- "https://www.justice.gov/criminal-fraud/cares-act-fraud"

doj_pppfraud_cases <- doj_pppfraud_url %>%
  read_html() %>% 
  html_elements('ul li b') %>% 
  html_text() %>% 
  as_tibble()


# display the html below
doj_pppfraud_cases
```

Q2. In how many individual judicial districts has the U.S. Department of Justice filed cases against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud?  Note: an example of a judicial district is "Southern District of Florida". You will need to use rvest scrape the data on this single webpage into a dataframe.
A2.
There are 31 different judicial districts that the DOJ filed cases in.
```{r}
doj_pppfraud_url <- "https://www.justice.gov/criminal-fraud/cares-act-fraud"

doj_pppfraud_jurisdictions <- doj_pppfraud_url %>%
  read_html() %>% 
  html_elements('p b i') %>% 
  html_text() %>% 
  as_tibble()

  # display the html below
doj_pppfraud_jurisdictions
```

Q4. The website (https://www.justice.gov/criminal-fraud/cares-act-fraud) shows that the DOJ filed more cases in the Southern District of Florida than in any other district. One of those cases was filed against someone named "Diamond Blue Smith". Who is Smith, and what was he accused of, according to the criminal complaint? If you were an editor, would you have assigned a reporter to write a story about this case when the complaint was filed in court? Why or why not?
A4.

Smith, a Florida recording artist, was charged for participation in a scheme to use fraudulent loan applications in PPP loans. Smith's documents got him over a million dollars in fraudulent PPP loans. Using that money, he bought a Ferrari and other luxury purchases. I would assign a reporter for writing the story about this case because there's a level of extravagance that gives the story notoriety.

Q5. In what percentage of all judicial districts has the U.S. Department of Justice filed cases cases against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud? In answering this question, you should also produce a list of judicial districts where DOJ has NOT filed a case, according to this site.  Note: to answer this question, you will need to scrape a table of all district courts on this up-to-date Wikipedia page under the heading "Active Courts": https://en.wikipedia.org/wiki/List_of_United_States_district_and_territorial_courts  
A5.
33 percent of jurisdictions were ones that the DOJ files a case.

```{r}
doj_pppfraud_url <- "https://www.justice.gov/criminal-fraud/cares-act-fraud"

doj_pppfraud_jurisdictions <- doj_pppfraud_url %>%
  read_html() %>% 
  html_elements('p b i') %>% 
  html_text() %>% 
  as_tibble()


# display the html below
doj_pppfraud_jurisdictions

districts_wiki_url <- "https://en.wikipedia.org/wiki/List_of_United_States_district_and_territorial_courts"

districts_wiki <- districts_wiki_url %>%
  read_html() %>% 
  html_table()

districts_wiki <- districts_wiki[[3]]


nonactive_courts <- districts_wiki %>% anti_join(doj_pppfraud_jurisdictions, by=c("Region" = "value"))

percentage_count = (31/94)
percentage_count
```
Q6. What might explain why, according to this site, the DOJ has filed PPP-related fraud cases in less than half of the country's judicial districts?
A6.

It could be because of the laws in those areas, they could be more lax in these areas and thereby incentivize people to attempt fraud.

Q7. Which state had the most approved PPP loans per 100,000 population? [This web page](https://dwillis.github.io/jour472files/ppp_scraping_example/index.html) has links to 52 individual web pages, one for each state (plus Washington, D.C. and Puerto Rico). Each of those web pages contains a one-row html table that has the name of the state, the number of approved loans, and the 2019 population. Here's an example for [Alabama](https://dwillis.github.io/jour472files/ppp_scraping_example/states/alabama.html). You'll need to loop over the individual state urls, scrape each individual page and combine the information on each page into a single dataframe to answer this question. Don't forget to calculation the loans per 100,000.
A7.
South Dakota got the most loans per 100,000 people with 7379.531.
```{r}

ppp_state_data <- tibble()

ppp_overalldata_url <- "https://dwillis.github.io/jour472files/ppp_scraping_example/index.html"
ppp_overalldata <- ppp_overalldata_url %>% 
read_html() %>% 
html_table()

ppp_overalldata <- ppp_overalldata[[1]] %>% 
  filter(state_page_link != "Puerto Rico") %>% 
  filter(state_page_link != "District of Columbia")

for(row_number in 1:nrow(ppp_overalldata)) {
      each_row_df <- ppp_overalldata %>%
        slice(row_number)
      
      url <- each_row_df$url
      
      ppp_state_data_sep <- url %>%
        read_html() %>%
        html_table()
      
      
        ppp_state_data_sep <- ppp_state_data_sep[[1]]
        
      ppp_state_data <- ppp_state_data %>%
        bind_rows(ppp_state_data_sep)
      
}
```

``` {r}
      ppp_state_data <- ppp_state_data %>% 
        mutate(ppp_per100k = (total_ppp_loans/population)*100000) %>% 
        arrange(desc(ppp_per100k))

ppp_state_data
```
